### 2014-07-23 markdowns/test.Rmd

---
title: "test"
author: "Court Campany"
date: "Tuesday, July 22, 2014"
output: html_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r loads functions packages and plot objects, echo=FALSE}
#source functions, packages, anbd plot objects
library(knitr)
opts_knit$set(root.dir=normalizePath('../'))
source("functions and packages/startscripts.R")

```