### 2014-09-03 markdowns/master.rmd
change to C gain model and master.Rmd

---
title: "EUC_PVE"
author: "Court Campany"
date: "Tuesday, July 22, 2014"
output: html_document
---

```
These are the plots and relevatn statistics for the Eucalpytus tereticornis  
pot study.  6 total contatiner volumes and 'free'seedlings all grown under  
natural field conditions during the summer of 2012
```
```{r setup}
library(knitr)
opts_knit$set(root.dir = '../')
```
```
```{r loads functions packages and plot objects, echo=FALSE, include=FALSE}
#source functions, packages, anbd plot objects
source("functions and packages/startscripts.R")
```

seedling growth through time

```{r allometry panel plot}
source("master_scripts/allometryplotting.R")
```

seedling total leaf area through time

```{r leafareaplot}
source("master_scripts/leafarea_plotting.R")
```

root:shoot and leaf:fine root at harvest

```{r rootshootplots }
source("master_scripts/rootshoot_plotting.R")
```

asat and amax

```{r photosynthesisplotting}
source("master_scripts/A_plotting.R")
```

aci curves and jmax/vcmax

```{r acimeansplots }
source("master_scripts/aci_plotting.R")
```

### seedling predawn and midday water potentials

```{r waterpotentialplots}
source("clean scripts/seedling water potential.R")
```

Modelled C gain versus harvest seedling total carbon

```{r CgainvsCmass }
cgain<- read.csv("calculated data/euc_cgain.csv")
cgain$volume <- as.factor(cgain$volume)

windows()
with(cgain, plot(carbon_gain, totalC, pch=pchs[volume], col=volume))
abline(0,1)
abline(lm(totalC ~ carbon_gain, data=cgain), lty=5)
```

Soil N percent pre expeeriment and at harvest

```{r SoilNpercent}
source("master_scripts/soilCN_plotting.R")
```

Predicted responses of Amax on a mass basis to leaf starch and N levels
```{r AmassvsTNCandN}
source("master_scripts/photochem_predict.R")
```