### 2017-03-02 manuscript/editor_response2_tree_physiology.Rmd
maunscript comment edits

---
output:
  word_document:
    reference_docx: manuscriptstyle_pce2.docx
  pdf_document: default
  html_document:
    number_sections: yes
---

```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

Prof Daniel Epron
\
Editor
\
Tree Physiology

\
\
We are grateful for the opportunity to resubmit our manuscript. We thank the editor for providing insightful comments to improve our manuscript. We have responded in detail to each comment, which has improved the overall quality of the manuscript. Below, we have itemized our response to each comment from the editor.

\
- *Comment: 1. The abstract should be revised according to the main document. For example, there is no reduction in whole-plant C use efficiency (CUE) with container size, just a difference between container versus Free rooting.*

\
*Response*: We have corrected the CUE findings in the abstract to reference the decrease of CUE between Free seedlings and seedlings in containers, not of container size.

\
- *Comment: 2. Increasing rather than decreasing C sink activity is another approach that you did not mentioned in your introduction. It was done on rubber tree by comparing trees that are tapped for latex production with trees that are not tapped. You may have a look on the two following papers and see whether or not it can be included in your introduction.*

\
*Response*: Although both studies with rubber trees are extremely interesting in terms of the dymanics of storage carbohyrdate accumluation and mobilization, they lack empirical data for photosynthesis rates and only speculate about the impacts on growth. I have choosen to not mention them in the introduction, as they are mostly out of context with discusssion of the effects of sink limitation on physiological processes and growth.

\
- *Comment: 3. g1 is not generated using only gs values. Provide more details please.*

\
*Response*: We have corrected to this statement to include observed values of *g*~s~, *A*~sat (which includes C~a~ of 400) and vapor presssure deficit.

\
- *Comment: 4. The Q10 relation (eqn 2) is much older than Atkin 2005. Please remove this useless reference.*

\
*Response*: Reference to Atkin has been removed.

\
- *Comment: 5. L267: misusing punctuations*

\
*Response*: editorial change made

\
- *Comment: 6. L275: Recall us here what "An" is.*

\
*Response*:

\
- *Comment: 7. L294: seedling C biomass (add C for clarity)*

\
*Response*: editorial change made

\
- *Comment:  8. L249-294: be consistent when using "P" or "Pday"*

\
*Response*: *P* corrected to *P*~day~ throughout this entire subsection.

\
- *Comment: 9. L300: "Amass" should be introduced before*

\
*Response*: We have now introducted the term *A*~mass~ in this section as it had not been previously described.

\
- *Comment: 10. L321-322 :"differences were driven mainly by the largest container and the Free seedling treatments". This is not what we can see on Table 1, at least for leaf area. Can final height and diameter added in this table, with their stats?*

\
*Response*: We agree that this result cannot be seen from clearly from Table 1.  However, it can be seen easily from Figure 1. Instead of increasing the size of an already large table, I have referenced Figure 1 at then end of this statement for simplicity.

\
- *Comment:11. L370-371: g1 "was lowest in the Free seedling treatment and was marginally different across soil volume treatments". This is not so clear in Table 3. it seems that larger soil volume are more similar to Free than to smaller soil volume.*

\
*Response*: We have reworded the desciprtion of the findings with the g~1~ parameter. "The *g*~1~ parameter ... was highest in the smaller containers and decreased in larger containers and the Free seedling treatment"

\
- *Comment: 12. L386: replace "-" by "versus" or "vs" or anything else to avoid confusion*

\
*Response*: changed "-" to vs

\
- *Comment: 13. L430: uncorrect "Aat"*

\
*Response*: corrected to *A*~sat~

\
- *Comment: 14. L468-469: a reference is needed for "functional partitioning of ..."*

\
*Response*:

\
- *Comment: 15. L487 recall us again here what "An" is*

\
*Response*:

\
- *Comment: 17. L512: is "however" really needed?*

\
*Response*: removed "however"

\
- *Comment: 17. Fig3c: this is not plant total mass*

\
*Response*: